### 2016-08-01 Projects/DelCo_2012/Report/DelCo_2012.tex
Added final edits to CAFS report, including new refs

All of the experiments were conducted at full scale in compartments and structures that were representative of room sizes in a residential occupancy. Technical panel members whose departments routinely use CAF for structural fire fighting were invited and present at many of the fire suppression experiments. The experiments were planned during the workshop.  Three types of experiments examining spray density measures, gas layer cooling, and fire suppression were conducted in the fall of 2012. The data from these experiments were presented to the technical panel members. The fire suppression experiments were conducted with a fuel package that included upholstered furniture. After a review of the data, a decision was made to conduct an additional series of suppression experiments with a fuel package that was composed mainly of wood-based materials. The thinking was that the synthetic materials in the upholstered furniture burned only on the surface and were easily extinguished. wood-based fuels would char and present a deep seated fire challenge to the extinguishing agents.  After the wood-based fuel package experiments were completed in 2013, it was decided in discussions with members of the technical panel that other experimental conditions needed to be examined: room size, additional ventilation, and a manually operated handline. The issue of a larger fire compartment had also been raised during the workshop discussion, but the desire for replicate experiments and financial limitations removed the large floor plan experiments from the original plan. To accomplish the last series of experiments, additional funding was obtained from NIST, and the experiments were conducted in 2015. The fuel load contained both wood-based wall and ceiling finishes, pallets, as well as carpeting and upholstered furniture.

@@

The west structure was modified to add second story to test larger fires and remote suppression effects (see Figure~\ref{fig:delco_2story}).

@@

The outer walls of the bottom floor of the two-story structure were identical to that of the single-story structures described above but with an open floor plan. The interior dimensions were 6.1~m (20~ft) wide, 11~m (36~ft) long and 2.4~m (8~ft) high. Figure~\ref{fig:dimensioned_first_2story} shows the dimensions of the reconfigured first floor. A stairwell was built to connect the floors. Stairs started 1.6~m (5.25~ft) off the South wall with a width of 1.2~m (4~ft) off the East wall. The second story walls were wood framed with 51~mm (2~in) by 102~mm (4~in) studs. The studs were set to 0.40~m (16~in) centers. The interior  walls were protected by 16~mm (5/8~in) fire rated gypsum board, 16~mm (5/8~in) durarock board, and a second layer of 16~mm (5/8~in) fire rated gypsum board. The exterior walls were protected with 11~mm (7/16~in) oriented strand board and 8~mm (5/16~in) fiber cement lap siding. Figure~\ref{fig:dimensioned_second_2story} shows the dimensions of the second story of the modified west structure.

@@

The gas cooling tests included 4 bare-bead thermocouple arrays, a bi-directional probe with a thermocouple array, and a total heat flux gauge/radiometer pair. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions} provides the positions within the burn room and gas cooling room where the senors were located (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each of the bare-bead thermocouple arrays had 11 thermocouples. The top thermocouple was placed at the ceiling and each subsequent thermocouple was spaced 0.3~m (1~ft) apart with the bottom thermocouple being 3.05~m (10ft) below the ceiling. There were 6 velocity probes and solid thermocouples centered at the external doorway to the gas cooling room. The top probes were 0.15~m (0.5~ft) below the door soffit, the second probes were 0.3~m (1~ft), and the remaining 4 were spaced 0.3~m (1~ft) apart with the bottom probe being 1.52~m (5~ft) below the soffit. The total heat flux gauge/radiometer set were positioned 0.15~m (0.5~ft) off the ground and aimed to view the ceiling.

@@

The fire suppression testing in the single-story structure included 2 bare-bead thermocouple arrays, three bi-directional probe plus solid thermocouple arrays, 2 total heat flux gauge pairs, and 2 total heat flux gauge/radiometer pairs. Figure~\ref{fig:Fire_Suppression_Instrumentation_Dimensions} provides the positions within the burn room, hallway, and entrance foyer where the senors were located (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each bare-bead thermocouple array had 8 thermocouples. The top sensor was 0.03~m (1~in) below the ceiling and the remaining thermocouples were spaced 0.3~m (1~ft) apart with the bottom thermocouple being 2.13~m (7~ft) below the ceiling. The three bi-directional probe arrays had unique sensor locations. For the window array, there were four sensor pairs located 0.28~m (0.95~ft), 0.58~m (1.90~ft), 0.84~m (2.85~ft), and 1.16~m (3.8~ft) below the soffit. For the hallway array, the 7 sensor pairs started 0.3~m (1~ft) below the ceiling and were spaced every 0.3~m (1~ft) ending 2.13~m (7~ft) below the ceiling. The doorway array also had 7 sensor pairs, but the first sensors were located 0.15~m (0.5~ft) below the soffit. The remaining sensors were spaced every 0.3~m (1~ft) with the bottom pair 1.83~m (6~ft) below the doorway soffit. The total heat flux gauge/radiometer pairs were set to be 0.15~m (0.5~ft) off the ground and aimed to view the ceiling. The pairs of total heat flux gauges were set 1~m (3~ft) off the ground with one sensor ``looking'' vertical and the other horizontal.

@@

Fire suppression testing in the two-story structure included similar types of instrumentation but included more sensors. Three bare-bead thermocouple arrays and 2 bi-directional probe and thermocouples arrays were used in the first floor. Figure~\ref{fig:fire_supp_first_2story} shows the positions of the instrumentation (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each of the bare-bead arrays had 8 thermocouples with the same spacing as the single-story tests. The bi-directional probe and solid thermocouple arrays also had 8 sensors with the first probe 0.08~m (0.25~ft) below the soffit with the remaining probes at 0.34~m (1.1~ft), 0.61~m (2~ft), 0.88~m (2.9~ft), 1.15~m (3.7~ft), 1.42~m (4.7~ft), 1.68~m (5.5~ft), and 1.95~m (6.4~ft) below the soffit.

@@

The second story featured 6 bare-bead arrays and 3 bi-directional probe plus solid thermocouple arrays. The second floor also had three heat flux sensor pairs (one facing horizontal and one facing vertical) that were 1~m (3~ft) off the ground. Figure~\ref{fig:fire_supp_second_2story} shows the positions of the instrumentation (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). The bare-bead thermocouple arrays featured 8 sensors with the same spacing as the basement level. The bi-direction probe arrays at the south wall door and at the top of the stairwell were comprised of 8 sensor pairs with the same spacing as the basement level. The third bi-directional probe array (3~m (9~ft) off the North wall and 0.6~m (2~ft) off the East wall) has 8 sensor pairs with the same spacing at the bare-bead thermocouple arrays.

@@

Water flow rate was measured with a pressure and flow meter combination. The meter consists of a section of 6.35~cm (2.5~in) cast aluminum pipe with a 0--4.1~MPa (0--600~psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800~lpm (1250~gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate.  The manufacturer reports a $\pm$ 5~\% calibration expanded uncertainty for the flow sensor and $\pm$ 3~\% calibration expanded uncertainty for the pressure sensor~\cite{Akron:2009}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flow rate. The total expanded uncertainty is estimated at $\pm$ 10~\%.

@@

Wood pallets were used as the fuel source for the gas cooling experiments. The pallets were approximately 1.2~m (4.0~ft) by 1.0~m (3.3~ft) by 0.13~m (0.42~ft) thick and ranged in mass from 13.6~kg (29.9~lb) to 26.4~kg (58.1~lb) with an average of 18.4~kg (40.5~lbs). The initial fuel load consisted of 10 pallets, arranged in two stacks of five, as shown in Figure~\ref{fig:Burn_Building_Fuel_Load}. Approximately one half of a bale of excelsior, 13.0~kg (28.6~lb), was mixed with the pallets to aid with ignition. Figure~\ref{fig:Burn_Building_Fuel_Load} also provides dimensioned locations of the two stacks of pallets within the burn room.

@@

The fuel load for test numbers 1 through 8 consisted of two sleeper sofas, two cushioned chairs, carpet, padding, and wood paneling along the southeast corner walls. An image of the fuel arrangement is shown in Figure~\ref{fig:Furniture_Fuel_Load}. The couches and chairs were set up to mimic a typical seating area with one couch against the south wall and one couch against the east wall, each positioned 0.9~m (36~in) away from the southeast corner. A chair was placed adjacent to each couch. The specific locations can be found in the dimensioned schematic, Figure~\ref{fig:Furniture_Fuel_Load_Dimensions}. In total, this fuel package contains approximately 336~kg (739~lbs) of fuel consisting of wood, flexible polyurethane (PU), polyester (PE),and polypropylene (PP). Table~\ref{tab:Fire_Suppression_Fuel_Masses} provides a breakdown on the fuel composition for these experiments.

@@

In June 2015, two stacks of pallets were added to the fuel load consisting of three sofas; OSB lining on the ceiling and walls; and the polypropylene carpeting over polyurethane foam padding. Each stack had 10 pallets and an approximate mass of 175~kg (386~lbs). Hay from bales averaging a mass of 12~kg (26~lbs) was added for ignition purposes.

@@

The quality of the fit of the red dashed line to the scatter in gas cooling data is represented by the coefficient of determination (R$^2$ on Figure~\ref{fig:combined_all}). The coefficient of determination provides a way to quantify how significant the variation of the data is compared the the least squared error line. Here, the higher the value of R$^2$, the better the ordinary least square line describes the relationship between water and CAF temperature difference comparisons.

@@

For the 12 middle position replicate comparisons (Figure~\ref{fig:CAFS_Water_mid}), the results (via the dashed red line) show that water has a slight advantage in gas cooling relative to CAF. At the back position (Figure~\ref{fig:CAFS_Water_full}), for the 3 replicate comparisons, the least squared error line falls within the experimental uncertainty. This indicates an inability to distinguish between the two suppression agents with statistical confidence.

To further analyze the data, the temperature differences can be separated by each of the three thermocouple arrays in the fire room. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2} shows the placement of three thermocouple arrays (TC A1, TC A2, and TC A3).  Ideally, the thermocouple is providing an estimate of the temperature of the gas that it is surrounded by.  However, we must remember that the thermocouple is actually providing us with its own temperature.  So if a droplet of water or some CAF cling to a thermocouple, the temperature provided by that thermocouple will be different than the surrounding gas temperature.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/PDFs/West_Structure/DelCo_2012_West_Structure_Instrumentation2}
	\caption{Schematic of  the Gas Cooling Experiments}
	\label{fig:Gas_Cooling_Instrumentation_Dimensions2}
\end{figure}
Based on the monitor position (arrow at the doorway in Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2}), thermocouple arrays A1 and A2 had more direct impact from the suppression agent compared to thermocouple array A3. As an example, Figure~\ref{fig:Burn_Building_Test_18b} shows the distribution of the agent within the burn building from a spray density test for a straight stream with CAF at the mid position compared to the positions of the thermocouple trees.
\begin{figure}[!ht]
	\includegraphics[width=.7\columnwidth]{../Figures/Bars/BB18b}
	\caption{Gas Cooling Spray Density Test 18, CAF, Straight Stream}
	\label{fig:Burn_Building_Test_18b}
\end{figure}

There is a difference in the amount of direct impact from the suppression agent at the three spatial locations. Therefore, the maximum temperature differences for each thermocouple on each of the three arrays should also be analyzed separately. Figures~\ref{fig:CAFS_Water_A1_all}, \ref{fig:CAFS_Water_A2_all}, and \ref{fig:CAFS_Water_A3_all} show the temperature cooling scatter for thermocouple trees A1, A2, and A3, respectively. Note that thermocouple array A3, the array which has the least amount of direct impact from the suppression agents, has the closest fit compared to the solid black line that represents uniform impact for both suppression agents. The fit lines for A1 and A3 also fall within the experimental uncertainty, meaning that differences in the impact of CAF and water are indistinguishable. Thermocouple tree A2, which gets the most significant direct impact from the suppression agents, shows a bias towards water being more effective than CAF on cooling the thermocouples.

@@

The thermocouple array data can also also be decomposed as a function of nozzle position: 12 middle position experimental pairs and 3 back experimental pairs. Figures~\ref{fig:CAFS_Water_A1_mid}, \ref{fig:CAFS_Water_A2_mid}, and \ref{fig:CAFS_Water_A3_mid} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 12 replicate tests with the monitor nozzle in the middle position. Figures~\ref{fig:CAFS_Water_A1_back}, \ref{fig:CAFS_Water_A2_back}, and \ref{fig:CAFS_Water_A3_back} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 3 replicate tests with the monitor nozzle in the back position. Again note that the best fit occurs with the data from thermocouple array 3 for both nozzle positions.  This analysis of the data results in the same conclusions at the analysis from the Mitchell paper \cite{Mitchell:1}, the gas cooling performance of CAF, under these conditions, is similar to water.

@@

There were 12 suppression tests conducted in the single-story purpose-built concrete structure. Eight of the experiments used a fuel load comprised of typical household furnishings that included wooden wall paneling, two sofas, two overstuffed chairs, carpet, and padding (Section~\ref{sec:fire_suppression_furniture_fuel}). The other four experiments used wood pallet fuel load in a space with the ceiling and walls composed of wood-based materials (Section~\ref{sec:fire_suppression_pallet_fuel}). Ignition occurred in the rear of the single-story structures and the available ventilation was through the front doorway and for eight of the experiments a rear window was ``opened''to provide additional ventilation (see Table~\ref{tab:Test_Descriptions}). The fires were intended to replicate post-flashover residential room and contents fires. Hose lines were pre-deployed with fixed nozzles set in two locations: one in the hallway on the approach to the burn room (hallway nozzle) and one just inside the burn room to the right side of the hallway (door nozzle). Each nozzle flowed 15~s of water or compressed air foam depending on the test. The hallway nozzle was intended to simulate a crew advancing into a structure towards the burn room. Suppression from this location would be in the form of gas cooling as the stream was not aimed at the seat of the fire (indirect interior attack).  After the hallway nozzle was shutdown, there was a 15~s time period to simulate the crew advancing into the doorway of the fire room. The door nozzle was positioned just to the side of the entry door.  It was also opened and flowed for approximately 15~s. The door nozzle was aimed in the direction of the seat of the fire (direct interior attack). Table~\ref{tab:Test_Descriptions} provides an overview of the suppression tests conducted.

@@

Table~\ref{tab:Test_Results} provides the impact of the two nozzle applications for the water and CAF tests. The table provides change in temperature 0.3~m (1~ft) below the ceiling and 0.3~m (1~ft) above the floor as well as change in heat flux 0.15~m (6~in) above the floor at the measurement location closest to the fire. These values reported in the table are immediately before and after the 15~s application period to assess the application agent, stream, and location on interior conditions. Remember that there was a 15~s gap between the hallway nozzle flow and the door nozzle flow. The amount of temperature change during the period in between nozzle flows is shown by comparing the temperature at the end of the hallway flow period with the temperature at the beginning of the room flow period.

>>>>

All of the experiments were conducted at full scale in compartments and structures that were representative of room sizes in a residential occupancy. Technical panel members whose departments routinely use CAF for structural fire fighting were invited and present at many of the fire suppression experiments. The experiments were planned during the workshop.  Three types of experiments examining spray density measures, gas layer cooling, and fire suppression were conducted in the fall of 2012. The data from these experiments were presented to the technical panel members. The fire suppression experiments were conducted with a fuel package that included upholstered furniture. After a review of the data, a decision was made to conduct an additional series of suppression experiments with a fuel package that was composed mainly of wood-based materials. The thinking was that the synthetic materials in the upholstered furniture burned only on the surface and were easily extinguished. wood-based fuels would char and present a deep seated fire challenge to the extinguishing agents.  After the wood-based fuel package experiments were completed in 2013, it was decided in discussions with members of the technical panel that other experimental conditions needed to be examined: room size, additional ventilation, and a manually operated handline. The issue of a larger fire compartment had also been raised during the workshop discussion, but the desire for replicate experiments and financial limitations removed the large floor plan experiments from the original plan. To accomplish the last series of experiments, additional funding was obtained from NIST, and the experiments were conducted in 2015. The fuel load contained both wood-based wall and ceiling finishes, pallets, carpeting, and upholstered furniture.

@@

The west structure was modified to add a second story to test larger fires and remote suppression effects (see Figure~\ref{fig:delco_2story}).

@@

The outer walls of the bottom floor of the two-story structure were identical to that of the single-story structures described above but with an open floor plan. The interior dimensions were 6.1~m (20~ft) wide, 11~m (36~ft) long and 2.4~m (8~ft) high. Figure~\ref{fig:dimensioned_first_2story} shows the dimensions of the reconfigured first floor. A stairwell was built to connect the floors. Stairs started 1.6~m (5.25~ft) off the south wall with a width of 1.2~m (4~ft) off the east wall. The second story walls were wood framed with 51~mm (2~in) by 102~mm (4~in) studs. The studs were set to 0.40~m (16~in) centers. The interior  walls were protected by 16~mm (0.63~in) fire rated gypsum board, 16~mm (0.63~in) durarock board, and a second layer of 16~mm (0.63~in) fire rated gypsum board. The exterior walls were protected with 11~mm (0.31~in) oriented strand board and 8~mm (0.31~in) fiber cement lap siding. Figure~\ref{fig:dimensioned_second_2story} shows the dimensions of the second story of the modified west structure.

@@

The gas cooling tests included four bare-bead thermocouple arrays, a bi-directional probe with a thermocouple array, and a total heat flux gauge/radiometer pair. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions} provides the positions within the burn room and gas cooling room where the sensors were located (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each of the bare-bead thermocouple arrays had 11 thermocouples. The top thermocouple was placed at the ceiling and each subsequent thermocouple was spaced 0.3~m (1~ft) apart with the bottom thermocouple being 3.05~m (10ft) below the ceiling. There were 6 velocity probes and solid thermocouples centered at the external doorway to the gas cooling room. The top probes were 0.15~m (0.5~ft) below the door soffit, the second probes were 0.3~m (1~ft) below the soffit, and the remaining 4 were spaced 0.3~m (1~ft) apart with the bottom probe being 1.52~m (5~ft) below the soffit. The total heat flux gauge/radiometer set were positioned 0.15~m (0.5~ft) off the ground and aimed to view the ceiling.

@@

The fire suppression testing in the single-story structure included 2 bare-bead thermocouple arrays, three bi-directional probe plus solid thermocouple arrays, 2 total heat flux gauge pairs, and 2 total heat flux gauge/radiometer pairs. Figure~\ref{fig:Fire_Suppression_Instrumentation_Dimensions} provides the positions within the burn room, hallway, and entrance foyer where the senors were located (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each bare-bead thermocouple array had 8 thermocouples. The top sensor was 0.03~m (1~in) below the ceiling and the remaining thermocouples were spaced 0.3~m (1~ft) apart with the bottom thermocouple being 2.13~m (7~ft) below the ceiling. The three bi-directional probe arrays had unique sensor locations. For the window array, there were four sensor pairs located 0.28~m (0.95~ft), 0.58~m (1.90~ft), 0.84~m (2.85~ft), and 1.16~m (3.8~ft) below the soffit. For the hallway array, the 7 sensor pairs started 0.3~m (1~ft) below the ceiling and were spaced every 0.3~m (1~ft) ending 2.13~m (7~ft) below the ceiling. The doorway array also had 7 sensor pairs, but the first sensors were located 0.15~m (0.5~ft) below the soffit. The remaining sensors were spaced every 0.3~m (1~ft) with the bottom pair 1.83~m (6~ft) below the doorway soffit. The total heat flux gauge/radiometer pairs were set to be 0.15~m (0.5~ft) off the ground and aimed to view the ceiling. The pairs of total heat flux gauges were set 1~m (3~ft) off the ground with one sensor ``looking'' vertical at the ceiling and the other horizontal.

@@

Fire suppression testing in the two-story structure included similar types of instrumentation but included more sensors. Three bare-bead thermocouple arrays and 2 bi-directional probe and thermocouples arrays were used in the first floor. Figure~\ref{fig:fire_supp_first_2story} shows the positions of the instrumentation (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). Each of the bare-bead arrays had 8 thermocouples with the same spacing as the single-story tests. The bi-directional probe and solid thermocouple arrays also had 8 sensors with the first probe 0.08~m (0.25~ft) below the soffit and the remaining probes at 0.34~m (1.1~ft), 0.61~m (2~ft), 0.88~m (2.9~ft), 1.15~m (3.7~ft), 1.42~m (4.7~ft), 1.68~m (5.5~ft), and 1.95~m (6.4~ft) below the soffit.

@@

The second story featured six bare-bead arrays and three bi-directional probe plus solid thermocouple arrays. The second floor also had three heat flux sensor pairs (one facing horizontal and one facing vertical) that were 1~m (3~ft) off the ground. Figure~\ref{fig:fire_supp_second_2story} shows the positions of the instrumentation (see Figure~\ref{fig:Instrumentation_Legend} for reference on the symbols). The bare-bead thermocouple arrays featured 8 sensors with the same spacing as the basement level. The bi-directional probe arrays at the south wall door and at the top of the stairwell were comprised of 8 sensor pairs with the same spacing as the basement level. The third bi-directional probe array (3~m (9~ft) off the north wall and 0.6~m (2~ft) off the east wall) has 8 sensor pairs with the same spacing at the bare-bead thermocouple arrays.

@@

Water flow rate was measured with a pressure and flow meter combination. The meter consists of a section of 6.35~cm (2.5~in) cast aluminum pipe with a 0--4.1~MPa (0--600~psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800~l/min (1250~gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate.  The manufacturer reports a $\pm$ 5~\% calibration expanded uncertainty for the flow sensor and $\pm$ 3~\% calibration expanded uncertainty for the pressure sensor~\cite{Akron:2009}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flow rate. The total expanded uncertainty is estimated at $\pm$ 10~\%.

@@

Wood pallets were used as the fuel source for the gas cooling experiments. The pallets were approximately 1.2~m (4.0~ft) by 1.0~m (3.3~ft) by 0.13~m (0.42~ft) thick and ranged in mass from 13.6~kg (29.9~lb) to 26.4~kg (58.1~lb) with an average of 18.4~kg (40.5~lb). The initial fuel load consisted of 10 pallets, arranged in two stacks of five, as shown in Figure~\ref{fig:Burn_Building_Fuel_Load}. Approximately one half of a bale of excelsior, 13.0~kg (28.6~lb), was mixed with the pallets to aid with ignition. Figure~\ref{fig:Burn_Building_Fuel_Load} also provides dimensioned locations of the two stacks of pallets within the burn room.

@@

The fuel load for test numbers 1 through 8 consisted of two sleeper sofas, two cushioned chairs, carpet, padding, and wood paneling along the southeast corner walls. An image of the fuel arrangement is shown in Figure~\ref{fig:Furniture_Fuel_Load}. The couches and chairs were set up to mimic a typical seating area with one couch against the south wall and one couch against the east wall, each positioned 0.9~m (36~in) away from the southeast corner. A chair was placed adjacent to each couch. The specific locations can be found in the dimensioned schematic, Figure~\ref{fig:Furniture_Fuel_Load_Dimensions}. In total, this fuel package contains approximately 336~kg (739~lb) of fuel consisting of wood, flexible polyurethane (PU), polyester (PE),and polypropylene (PP). Table~\ref{tab:Fire_Suppression_Fuel_Masses} provides a breakdown on the fuel composition for these experiments.

@@

In June 2015, two stacks of pallets were added to the fuel load consisting of three sofas; OSB lining on the ceiling and walls; and the polypropylene carpeting over polyurethane foam padding. Each stack had 10 pallets and an approximate mass of 175~kg (386~lb). Hay from bales averaging a mass of 12~kg (26~lb) was added for ignition purposes.

@@

The quality of the fit of the red dashed line to the scatter in gas cooling data is represented by the coefficient of determination (R$^2$ in Figure~\ref{fig:combined_all}). The coefficient of determination provides a way to quantify how significant the variation of the data is compared to the least squared error line. Here, the higher the value of R$^2$, the better the ordinary least square line describes the relationship between water and CAF temperature difference comparisons.

@@

For the 12 middle position replicate comparisons (Figure~\ref{fig:CAFS_Water_mid}), the results (via the dashed red line) show that water had a slight advantage in gas cooling relative to CAF. For the three back position replicate comparisons (Figure~\ref{fig:CAFS_Water_full}), the least squared error line falls within the experimental uncertainty. This indicates an inability to distinguish between the two suppression agents with statistical confidence.

To further analyze the data, the temperature differences can be separated by each of the three thermocouple arrays in the fire room. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2} shows the placement of three thermocouple arrays (TC A1, TC A2, and TC A3). Ideally, the thermocouple is providing an estimate of the temperature of the gas that surrounded it. However, we must remember that the thermocouple is actually providing us with its own temperature. So if a droplet of water or some CAF cling to a thermocouple, the temperature provided by that thermocouple will be different than the surrounding gas temperature.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/PDFs/West_Structure/DelCo_2012_West_Structure_Instrumentation2}
	\caption{Schematic of  the Gas Cooling Experiments}
	\label{fig:Gas_Cooling_Instrumentation_Dimensions2}
\end{figure}

Based on the monitor position (arrow at the doorway in Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2}), thermocouple arrays A1 and A2 had more direct impact from the suppression agent compared to thermocouple array A3. As an example, Figure~\ref{fig:Burn_Building_Test_18b} shows the distribution of the agent within the burn building from a spray density test for a straight stream with CAF at the middle position compared to the positions of the thermocouple trees.

\begin{figure}[!ht]
	\includegraphics[width=.7\columnwidth]{../Figures/Bars/BB18b}
	\caption{Gas Cooling Spray Density Test 18, CAF, Straight Stream}
	\label{fig:Burn_Building_Test_18b}
\end{figure}

There is a difference in the amount of direct impact from the suppression agent at the three spatial locations. Therefore, the maximum temperature differences for each thermocouple on each of the three arrays should also be analyzed separately. Figures~\ref{fig:CAFS_Water_A1_all}, \ref{fig:CAFS_Water_A2_all}, and \ref{fig:CAFS_Water_A3_all} show the temperature cooling scatter for thermocouple arrays A1, A2, and A3, respectively. Note that thermocouple array A3, the array that had the least amount of direct impact from the suppression agents, has the closest fit compared to the solid black line that represents uniform impact for both suppression agents. The fit lines for A1 and A3 also fall within the experimental uncertainty, meaning that differences in the impact of CAF and water are indistinguishable. Thermocouple array A2, which gets the most significant direct impact from the suppression agents, shows a bias towards water being more effective than CAF on cooling the thermocouples.

@@

The thermocouple array data can also be decomposed as a function of nozzle position: 12 middle position experimental pairs and 3 back experimental pairs. Figures~\ref{fig:CAFS_Water_A1_mid}, \ref{fig:CAFS_Water_A2_mid}, and \ref{fig:CAFS_Water_A3_mid} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 12 replicate tests with the monitor nozzle in the middle position. Figures~\ref{fig:CAFS_Water_A1_back}, \ref{fig:CAFS_Water_A2_back}, and \ref{fig:CAFS_Water_A3_back} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 3 replicate tests with the monitor nozzle in the back position. Again, note that the best fit occurs with the data from thermocouple array 3 for both nozzle positions. This analysis of the data results in the same conclusions as the analysis from the Mitchell paper \cite{Mitchell:1}: the gas cooling performance of CAF is similar to water under these conditions.

@@

There were 12 suppression tests conducted in the single-story, purpose-built concrete structure. Eight of the experiments used a fuel load comprised of typical household furnishings that included wooden wall paneling, two sofas, two overstuffed chairs, carpet, and padding (Section~\ref{sec:fire_suppression_furniture_fuel}). The other four experiments used a wood pallet fuel load in a space with the ceiling and walls composed of wood-based materials (Section~\ref{sec:fire_suppression_pallet_fuel}). Ignition occurred in the rear of the single-story structure, and the available ventilation was through the front doorway. For eight of the experiments, a rear window was opened to provide additional ventilation (see Table~\ref{tab:Test_Descriptions}). The fires were intended to replicate post-flashover residential room and contents fires. Hose lines were pre-deployed with fixed nozzles set in two locations: one in the hallway on the approach to the burn room (hallway nozzle) and one just inside the burn room to the right side of the hallway (door nozzle). Each nozzle flowed 15~s of water or CAF depending on the test. The hallway nozzle was intended to simulate a crew advancing into a structure towards the burn room. Suppression from this location would be in the form of gas cooling, as the stream was not aimed at the seat of the fire (indirect interior attack).  After the hallway nozzle was shutdown, there was a 15~s time period of no nozzle flow to simulate the crew advancing into the doorway of the fire room. The door nozzle was positioned just to the side of the entry door.  It was also opened and flowed for approximately 15~s. The door nozzle was aimed in the direction of the seat of the fire (direct interior attack). Table~\ref{tab:Test_Descriptions} provides an overview of the suppression tests conducted.

@@

Table~\ref{tab:Test_Results} provides the impact of the two nozzle applications for the water and CAF tests. The table lists the change in temperature 0.3~m (1~ft) below the ceiling and 0.3~m (1~ft) above the floor as well as the change in heat flux 0.15~m (6~in) above the floor at the measurement location closest to the fire. These values reported in the table are immediately before and after the 15~s application period to assess the application agent, stream, and location on interior conditions. Recall that there was a 15~s gap between the hallway nozzle flow and the door nozzle flow. The amount of temperature change during the period in between nozzle flows is shown by comparing the temperature at the end of the hallway nozzle flow period with the temperature at the beginning of the door nozzle flow period.
